\begin{abstract}
As the potential capabilities of autonomous systems expand, it is increasingly necessary to verify the safety properties which evolve in both spatial and temporal domains. Currently, many next generation autonomous systems for both military and civilian applications remain research projects because there is a lack of confidence that they can be used in safety critical applications. In this paper, we develop a framework for formally verifying the safety of autonomous systems operating in unstructured environments and in the presence of other unpredictable agents. Using autonomous vehicles as a running example, we present a decomposition of an autonomous systems' mission into scenarios and agents. The scenario modeled by instantiating agents to create a network of hierarchical communicating hybrid automata (HCHA). We propose that such scenarios clumsily coexist in multiple domains, which when considered together result in undecidable verification problems. We show a method which separates the modes of the HCHA of the target agent and solves a multiplicity of verification problems by applying the best hammer for each mode. In turn we show through a coupling of the problems using the richer HCHA representation as an oracle we are able to state verification results for a complex passing maneuver with nonlinear vehicle dynamics.
%Autonomous and semi-autonomous systems are now an accepted part of the economy and society, whether they are autonomous robots in shipping warehouses or autonomous appliance robots in the home.
%The success of these robots can be attributed, in part, to the fact that either they function in highly controlled environments where their interactions with other agents are structured, or their tasks are simple enough to accommodate unstructured and unpredictable environments.
%In this paper, we develop a framework for formally verifying the safety of autonomous systems operating in unstructured environments in the presence of other unpredictable agents. 
%Using autonomous vehicles as a running example, 
%we present a decomposition of an autonomous system's mission into scenarios and agents, and a rich representation of these two elements in terms of hierarchical communicating hybrid automata. 
%Our goal is to formally verify that the behavior of the autonomous system is safe in a given scenario, which possibly involves other unpredictable (nondeterministic) agents.
%Therefore, we include in our framework a translator from this rich representation into the formalisms of various formal verification tools, to enable the use of the most appropriate tool for a given verification task. 
%The fact that multiple tools are needed is illustrated with the verification of the safety of a lane change maneuver, which requires the use of a timed automaton verification tool, and a rechability tool.
%Another goal is to allow control engineers to use these formal tools as part of their design process. 
%Therefore, we develop a scenario authoring tool, and a domain-specific language with formal semantics, for enabling rapid prototyping and verification of controllers' designs.
%


\end{abstract}