\section{Case study}
\label{caseStudy}
\subsection{Scenario Description}
Passing manuever is scenario on 2 lane road, target vehicle, 1 environmental vehicle.
\subsection{Target Vehicle Model}
\begin{enumerate}
	\item Target vehicle has non linear dynamics described by bicycle model.
	\item Trajectories are defined for right to left lane switch and left to right lane switch.
	\item Hybrid system modes are described by drive forward and lane change.
	\begin{enumerate}
		\item Lane change is hierarchical and has sub-modes, right to left, straight, and left to right.
		\item Each submode has same dynamics but implements a different motion primitive.
	\end{enumerate}
\end{enumerate}
\subsection{Environmental Vehicle Model}
Environmental vehicle has simple 2D dynamics, no specific controller, non-determinism describes state evolution.
\begin{enumerate}
	\item Environmental vehicle must drive forwards.
	\item Environmental vehicle must not change lanes if other lane is occupied.
	\item At every step the environmental vehicle can apply +/- some specified level of acceleration.
	\item At every step environmental vehicle can switch lanes if condition 2 is not violated.
\end{enumerate}
\subsection{dReal Model}
\begin{enumerate}
	\item 1 or 2 modes depending on lane change or passing.
	\item Dynamics are non linear bicycle model (insert figure)
	\item Plant controller is tracking controller which follows trajectory.
\end{enumerate}
\subsection{Controller with Partial Dynamics}
A finite transition system describing the passing manuever is given in UPPAAL (or NuSMV).