\section{outline}
\label{outline}
\todo[inline]{this section will be removed from final paper, of course}
Introduction
\begin{itemize}
	\item The benefits of autonomous systems.
	
	\item The acceptance of autonomous systems in the economy and our daily lives.
	
	\item Autonomous vehicles as a running example. What are the "real-world" challenges? (e.g. lane change in some detail, then just list others like pedestrians, other drivers,etc)
	
	\item What are the technical challenges facing the development of autonomous vehicles? (e.g. properties with priorities)
	
	\item How are these challenges addressed in today's technology and literature?
	
	\item What remains to be done?
	
	\item Which piece of the puzzle are we addressing in this paper? Illustrate with the running example of a lane change maneuver.
	
	\item How are we addressing it?
	
	\item The need for formal methods.
\end{itemize}



Modeling framework
\begin{itemize}
	\item Scenarios and agents (don't dwell too long on this, just need to mention that we are verifying one scenario in this paper, but we have a plan for how this fits with overall mission verification)
	
	\item HCHA: because we need a representation rich enough to express on it different types of properties, which apply to different components of the complex autonomous vehicle
	
	\item Specification languages: should be suited to the verification goal. Here, reachability and LTL

\end{itemize}

Lane change in the formalisms.
\begin{itemize}	
	\item Formal model of lane change in terms of HCHA for system, LTL for scenario goal, and reach sets for safety. Controller is a discrete controller in grid world.
	
	\item Lane change requirements can be verified with timed automata, so need to translate HCHA. Give translation.
	
	\item How results are mapped back to HCHA
	
\end{itemize}


Experiment
\begin{itemize}
	\item Experimental setup: simulation or actual little vehicles? UPPAAL, dReach, make it available online
	
	\item results
	
\end{itemize}

Future work
\begin{itemize}
	\item Tool chain: from design entry to formal DSL (implementing HCHA) to design files for relevant tools
	\item priority of requirements?
\end{itemize}